Carrier phase and amplitude estimation of $M$-ary phase shift keying using pilots and data

### 2012-05-10 paper.tex
Fixed silly error in the plots and resized them so that they fit 4 to a page.  More edits to proofs.

\begin{algorithm} \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's least squares estimator}
\DontPrintSemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$i \in D$ \nllabel{alg_loop_setup}}{
$\phi = \angle{y_i}$ \;
$u = \frac{2\pi}{M}\round{\frac{M}{2\pi}\phi} $ \;
$z_i = \phi -  u $ \;
$g_i = y_i e^{-j u}$ \;
}
$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
$\hat{a} = \frac{1}{L} Y$ \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
$Y = Y + (e^{j2\pi/M} - 1) g_{\sigma(k)}$ \;
$g_{\sigma(k)} = g_{\sigma(k)} e^{j2\pi/M} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}

@@

\begin{lemma}\label{lem:Zconunif}
Let $G_L(\lambda) = \re(Z(\theta)$.  Then
\[
\prob\{\abs{G_L(\lambda) - G(\lambda)} > \epsilon \} = O(e^{-\epsilon^2 L}).
\]
\end{lemma}
\begin{IEEEproof}
\end{IEEEproof}

\begin{IEEEproof} (Theorem~\ref{thm:normality})
The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.  This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the minimiser $\lambda}_L$. So
\[
G_L^\prime(\lambda_L) = 0 = \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\]
Write
\[
0 = G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) + \prob G_L^\prime(\lambda_L)
\]
where
\[
\prob G^\prime_L = \sum_{i \in P}
\]
\end{IEEEproof}

\begin{lemma}~\label{lem:diffatlambdaL}
The derivative of $G_L$ existsm, and is equal to zero, at $\lambda_L$.  That is,
\[
G_L^\prime(\lambda_L) = \frac{d G_L}{d \lambda}(\lambda_L) = 0.
\]
\end{lemma}

@@

%BPSK plots
\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotM2-1.mps}
		\caption{Amplitude error for BPSK}
		\label{fig:plotamp}
\end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphase}
\end{figure*}

@@

%QPSK plots
\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotM4-1.mps}
		\caption{Amplitude error for QPSK}
		\label{fig:plotamp}
\end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotM4-2.mps}
		\caption{Phase error for QPSK}
		\label{fig:plotphase}
\end{figure*}

% \begin{figure*}[tp]
% 	\centering
%         \includegraphics{code/data/plotM4-3.mps}
% 		\caption{Complex gain error for QPSK}
% 		\label{fig:plotcamp}
% \end{figure*}

%8PSK plots
\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotM8-1.mps}
		\caption{Amplitude error for 8PSK}
		\label{fig:plotamp}
\end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotM8-2.mps}
		\caption{Phase error for 8PSK}
		\label{fig:plotphase}
\end{figure*}

% \begin{figure*}[tp]
% 	\centering
%         \includegraphics{code/data/plotM8-3.mps}
% 		\caption{Complex gain error for 8PSK}
% 		\label{fig:plotcamp}
% \end{figure*}

@@

%BPSK plots
 \begin{figure*}[tp]
 	\centering
 		\includegraphics{code/data/plotM2-1.mps}
 		\caption{Amplitude error for BPSK}
 		\label{fig:plotamp}
 \end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotncM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphase}
\end{figure*}

%QPSK plots
 \begin{figure*}[tp]
 	\centering
 		\includegraphics{code/data/plotM4-1.mps}
 		\caption{Amplitude error for QPSK}
 		\label{fig:plotamp}
 \end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotncM4-2.mps}
		\caption{Phase error for QPSK}
		\label{fig:plotphase}
\end{figure*}

 % \begin{figure*}[tp]
 % 	\centering
 %         \includegraphics{code/data/plotM4-3.mps}
 % 		\caption{Complex gain error for QPSK}
 % 		\label{fig:plotcamp}
 % \end{figure*}

%8PSK plots
 \begin{figure*}[tp]
 	\centering
 		\includegraphics{code/data/plotM8-1.mps}
 		\caption{Amplitude error for 8PSK}
 		\label{fig:plotamp}
 \end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/data/plotncM8-2.mps}
		\caption{Phase error for 8PSK}
		\label{fig:plotphase}
\end{figure*}

% \begin{figure*}[tp]
% 	\centering
%         \includegraphics{code/data/plotM8-3.mps}
% 		\caption{Complex gain error for 8PSK}
% 		\label{fig:plotcamp}
% \end{figure*}

>>>>

\begin{algorithm} \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's least squares estimator}
\DontPrintSemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$i \in D$ \nllabel{alg_loop_setup}}{
$\phi = \angle{y_i}$ \;
$u = \round{\phi} $ \;
$z_i = \phi -  u $ \;
$g_i = y_i e^{-j u}$ \;
}
$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
$\hat{a} = \frac{1}{L} Y$ \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
$\eta = e^{j2\pi/M} - 1$ \;
\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
$Y = Y + \eta g_{\sigma(k)}$ \;
$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}

@@

\begin{lemma}\label{lem:Zconunif}
Let $G_L(\lambda) = \re(Z(\theta))$.  Then
\[
\prob\{\abs{G_L(\lambda) - G(\lambda)} > \epsilon \} = O(e^{-\epsilon^2 L}).
\]
\end{lemma}
\begin{IEEEproof}
\end{IEEEproof}

\begin{IEEEproof} (Theorem~\ref{thm:normality})
The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.  This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the minimiser $\lambda_L$. So
\[
G_L^\prime(\lambda_L) = 0 = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\]
Write
\[
0 = G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) + \prob G_L^\prime(\lambda_L)
\]
where
\[
\prob G^\prime_L(\lambda) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda),
\]
and where,
\begin{align*}
k_1(x) &= \prob R_1 \sin(x + \Phi_1) \;\;\; \text{and}, \\
k_2(x) &=  \prob R_1 \sin\sfracpart{x + \Phi_1}.
\end{align*}
Lemma~\ref{lem:GprimeLlambdaLconv} shows that
\[
\sqrt{L} \prob G^\prime_L(\lambda_L) = \sqrt{L} \lambda_L( G(0)  + o_P(1) )
\]
and Lemma~\ref{lem:empiricprocc} shows that
\[ 
\sqrt{L} \left( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \right) = \sqrt{L} G_L'(0) + \sqrt{L}\lambda_L( ? +  o_P(1) )
\]
It follows that
\[
\sqrt{\lambda_L} = o_P(1) + \frac{\sqrt{L}G_L'(0)}{G(0)  + o_P(1)}.
\]
Observe that 
\[
\sqrt{L} G_L'(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\]
It follows from the standard central limit that the distribution of
\[
\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{P}}{L}} \frac{1}{\sqrt{\abs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
\]
converges to the normal with mean $k_1(0) = 0$ and variance 
\[
p \prob R_1^2 \sin^2(\Phi_1).
\]
Similarly, the distribution of 
\[
\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{D}}{L}} \frac{1}{\sqrt{\abs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
\]
converges to the normal with mean $k_2(0) = 0$ and variance
\[
d \prob R_1^2 \sin^2\fracpart{\Phi_1}.
\]
It follows that the distribution of $\sqrt{L}\lambda_L$ converges to the normal with mean zero and variance
\[
\frac{p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}}{G(0)^2}.
\]
\end{IEEEproof}

\begin{lemma}~\label{lem:diffatlambdaL}
The derivative of $G_L$ exists, and is equal to zero, at $\lambda_L$.  That is,
\[
G_L^\prime(\lambda_L) = \frac{d G_L}{d \lambda}(\lambda_L) = 0.
\]
\end{lemma}

\begin{lemma}\label{lem:GprimeLlambdaLconv}
\[
\prob G^\prime_L(\lambda_L) = \lambda_L( G(0)  + o_P(1) )
\]
\end{lemma}
\begin{IEEEproof}
We have (BLERG) that $k_1$ and $k_2$ are continuous and differentiable and that $k_1(0) = k_2(0) = 0$, so by a first order Taylor expansion
\[
\prob G^\prime_L(\lambda_L) = \prob G^\prime_L(0) + \lambda_L( \prob G^\prime_L(\lambda_L)  + o_P(1) )
\]
\end{IEEEproof}

\begin{lemma}\label{lem:empiricprocc} BLERG
\[ 
\sqrt{L} \left( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \right) = \sqrt{L} G_L'(0) + \sqrt{L}\lambda_L( ? +  o_P(1) )
\]
\end{lemma}
\begin{IEEEproof}
\end{IEEEproof}

@@

%BPSK plots
\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
		\caption{Amplitude error for BPSK}
		\label{fig:plotamp}
\end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphase}
\end{figure}

@@

%QPSK plots
\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
		\caption{Amplitude error for QPSK}
		\label{fig:plotamp}
\end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
		\caption{Phase error for QPSK}
		\label{fig:plotphase}
\end{figure}

% \begin{figure}[tp]
% 	\centering
%         \includegraphics{code/data/plotM4-3.mps}
% 		\caption{Complex gain error for QPSK}
% 		\label{fig:plotcamp}
% \end{figure}

%8PSK plots
\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
		\caption{Amplitude error for 8PSK}
		\label{fig:plotamp}
\end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
		\caption{Phase error for 8PSK}
		\label{fig:plotphase}
\end{figure}

% \begin{figure}[tp]
% 	\centering
%         \includegraphics{code/data/plotM8-3.mps}
% 		\caption{Complex gain error for 8PSK}
% 		\label{fig:plotcamp}
% \end{figure}

@@

%BPSK plots
 \begin{figure}[tp]
 	\centering
 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
 		\caption{Amplitude error for BPSK}
 		\label{fig:plotamp}
 \end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphase}
\end{figure}

% \begin{figure}[tp]
% 	\centering
%         \includegraphics{code/data/plotM2-3.mps}
% 		\caption{Complex gain error for BPSK}
% 		\label{fig:plotcamp}
% \end{figure}

%QPSK plots
 \begin{figure}[tp]
 	\centering
 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
 		\caption{Amplitude error for QPSK}
 		\label{fig:plotamp}
 \end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
		\caption{Phase error for QPSK}
		\label{fig:plotphase}
\end{figure}

 % \begin{figure}[tp]
 % 	\centering
 %         \includegraphics{code/data/plotM4-3.mps}
 % 		\caption{Complex gain error for QPSK}
 % 		\label{fig:plotcamp}
 % \end{figure}

%8PSK plots
 \begin{figure}[tp]
 	\centering
 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
 		\caption{Amplitude error for 8PSK}
 		\label{fig:plotamp}
 \end{figure}

\begin{figure}[tp]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
		\caption{Phase error for 8PSK}
		\label{fig:plotphase}
\end{figure}

% \begin{figure}[tp]
% 	\centering
%         \includegraphics{code/data/plotM8-3.mps}
% 		\caption{Complex gain error for 8PSK}
% 		\label{fig:plotcamp}
% \end{figure}
